Timeliness of security updates for Android and Chrome

### 2014-02-14 securityupdates.tex
Finish sentance in securityupdates
Add some missing citations and urls

\documentclass[conference,a4paper,twoside]{IEEEtran}
\usepackage[utf8x]{inputenc}
\usepackage{ucs}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{url}
\usepackage{hyperref}
\author{Daniel R.\ Thomas, Daniel Wagner, Alastair R.\ Beresford and Andrew Rice}

@@

\section{Introduction}
Android has \percArdMarketShare\ of the smartphone market~\cite{TODO} and while the core development is controlled by Google there are at least \daNumManufacturers\ manufacturers which make devices which run Android.
Many of these manufacturers customise the version of Android they ship and sometimes network operators (of which there are at least \daNumNetworkOperators) make further modifications.
Hence when Google produces an update to Android, the update may have to pass through the manufacturer and operator before reaching the user.
The manufacturer and user have no financial incentive to perform this work.
There is ongoing legal action to force some of them to do so~\cite{TODO}.

@@

Figure \ref{fig:proportioninsecure} shows the proportion of Android devices in the DeviceAnalyzer data which were running versions of Android known to root exploits at the time.

@@

Chrome uses various different sandboxing methods on different platforms such as Capsicum on BSD~\cite{TODO}

@@

We have published full details of the vulnerabilities listed below in an accompanying technical report~\cite{TODO} and summarise them in Table \ref{tab:andvulns}.
We also have a website\footnote{\url{TODO}} which we will keep up to date with future vulnerabilities.

@@

\subsection{Versions of Android running on devices}
We also need historical information on which versions of Android were running on devices along with other information about those devices which allows us to go into more detail and confirm the reliability of the data our results depend on.
The only suitable data which is available for this is the DeviceAnalyzer data~\cite{TODO}.
The DeviceAnalyzer data contains data from ??? devices with a total of ??? device days, it contains data for more than ??? months for ??? devices.
%TODO talk about the DeviceAnalyzer project a bit
Various different kinds of data\footnote{\url{TODO}} are collected by the DeviceAnalyzer Android app\footnote{\url{TODO}}.
Among these are the build string and API version for the version of Android currently running on the device each day.
The API/SDK version is a well defined integer, however it does not always change with new Android releases, particularly security bug fixes should not increase this value.
The build string is `The user-visible version string', fortunately most (99.9\%) devices in the data have a build string of the form `x.y.z random string' and so it is possible to extract the Android version number.

@@

Figure \ref{fig:os} shows the raw data on the number of devices in the DeviceAnalyzer data running different versions of Android each day.
It shows how old versions are gradually replaced by new ones, the long tail of devices which do not see updates to more recent versions and the fluctuation in the number of devices in the data.
\begin{figure}%[!b]
\centering
\includegraphics[width=\columnwidth]{figures/os}
\caption{Android versions in DeviceAnalyzer data over time}
\label{fig:os}
\end{figure}

Figure \ref{fig:vulnerabilities} shows which vulnerabilities devices are exposed to.
For each vulnerability it shows the proportion of devices exposed to that vulnerability and how that changes over time.
In July 2011 at the beginning of the DeviceAnalyzer data the Exploid and Levitator vulnerabilities both affect most Android devices, slowly these are fixed as updates roll out and devices are replaced until in January 2013 a much smaller proportion of devices are affected by known vulnerabilities.
However when in February 2013 the first APK signing vulnerability was found which affected all previous versions of Android and even in October 2013 most devices were still unfixed. %TODO put percentages on these things to quantify them
\begin{figure}%[!b]
\centering
\includegraphics[width=\columnwidth]{figures/vulnerabilities}
\caption{Proportion of devices exposed to each vulnerability with time (1.0 is 100\%)}
\label{fig:vulnerabilities}
\end{figure}

>>>>

\documentclass[conference,a4paper,twoside]{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{url}
\usepackage{hyperref}
\RequirePackage[date=terse,isbn=true,doi=true,url=false,maxbibnames=9,backref=false]{biblatex}
\addbibresource{securityupdates.bib}
\renewcommand{\bibfont}{\small}

\AtEveryBibitem{% Clean up the bibtex rather than editing it
 \clearname{editor} % remove editors
}

\author{Daniel R.\ Thomas, Daniel Wagner, Alastair R.\ Beresford and Andrew Rice}

@@

\section{Introduction}
Android has \percArdMarketShare\ of the smartphone market~\cite{TODO} and while the core development is controlled by Google there are at least \daNumManufacturers\ manufacturers which make devices which run Android.
Many of these manufacturers customise the version of Android they ship and sometimes network operators (of which there are at least \daNumNetworkOperators) make further modifications.
Hence when Google produces an update to Android, the update may have to pass through the manufacturer and operator before reaching the user.
The manufacturer and user have no financial incentive to perform this work.
There is ongoing legal action to force some of them to do so~\cite{Soghoian2013}.
CESG which advises the UK government on how to secure its computer system recommends picking Android devices from manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which manufacturers are better.

@@

Figure \ref{fig:proportioninsecure} shows the proportion of Android devices in the Device Analyzer data which were running versions of Android known to root exploits at the time.

@@

Chrome uses various different sandboxing methods on different platforms such as Capsicum on BSD~\cite{Watson2010} or using a SUID binary to create new namespaces and chroot on Linux~\cite{TODO}.

@@

We have published full details of the vulnerabilities listed below in an accompanying technical report~\cite{TODO} and summarise them in Table \ref{tab:andvulns}.
We also have a website\footnote{\url{http://androidvulnerabilities.org/}} which we will keep up to date with future vulnerabilities.

@@

\subsection{Versions of Android running on devices}
We also need historical information on which versions of Android were running on devices along with other information about those devices which allows us to go into more detail and confirm the reliability of the data our results depend on.
The only suitable data which is available for this is the Device Analyzer data~\cite{Wagner2013}.
The Device Analyzer data contains data from ??? devices with a total of ??? device days, it contains data for more than ??? months for ??? devices.
%TODO talk about the Device Analyzer project a bit
Various different kinds of data\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.htm}} are collected by the Device Analyzer Android app\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/}}.
Among these are the build string and API version for the version of Android currently running on the device each day.
The API/SDK version is a well defined integer, however it does not always change with new Android releases, particularly security bug fixes should not increase this value.
The build string is `The user-visible version string', fortunately most (99.9\%) devices in the data have a build string of the form `x.y.z random string' and so it is possible to extract the Android version number.

@@

Figure \ref{fig:os} shows the raw data on the number of devices in the Device Analyzer data running different versions of Android each day.
It shows how old versions are gradually replaced by new ones, the long tail of devices which do not see updates to more recent versions and the fluctuation in the number of devices in the data.
\begin{figure}%[!b]
\centering
\includegraphics[width=\columnwidth]{figures/os}
\caption{Android versions in Device Analyzer data over time}
\label{fig:os}
\end{figure}

Figure \ref{fig:vulnerabilities} shows which vulnerabilities devices are exposed to.
For each vulnerability it shows the proportion of devices exposed to that vulnerability and how that changes over time.
In July 2011 at the beginning of the Device Analyzer data the Exploid and Levitator vulnerabilities both affect most Android devices, slowly these are fixed as updates roll out and devices are replaced until in January 2013 a much smaller proportion of devices are affected by known vulnerabilities.
However when in February 2013 the first APK signing vulnerability was found which affected all previous versions of Android and even in October 2013 most devices were still unfixed. %TODO put percentages on these things to quantify them
\begin{figure}%[!b]
\centering
\includegraphics[width=\columnwidth]{figures/vulnerabilities}
\caption{Proportion of devices exposed to each vulnerability with time (1.0 is 100\%)}
\label{fig:vulnerabilities}
\end{figure}
